---
title: "MEM Academic Data, Will Rains, Babak Barghi, Alexander Rutten"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows

---

```{r, message=FALSE}
library(ggplot2)
library(flexdashboard)
library(tidyverse)
library(RColorBrewer)
library(scales)
theme_set(theme_minimal())
students <- read.csv("Students.csv",header = TRUE,sep = ";")
grades <- read.csv("Grades.csv",header = TRUE,sep = ";")
courses <- read.table("Courses.csv", sep = ";", dec=".", header = TRUE)
courses <- tibble(courses)
courses <- courses %>%
  mutate(Course_ID = gsub(" ", "", Course_ID))
admissions <- read_csv2("Admissions.csv")
grades <- grades %>% filter(!is.na(Grade))

colnames(grades) <- c("Student_ID","Course_ID","Grade","Year")
```
Admissions
=====================================  

Column {data-height=500}
-----------------------------------------------------------------------

### AdmissionsByYear

```{r}
admissions %>% pivot_longer(-Enrolment, names_to = "Condition", 
                          values_to = "count") %>%
ggplot(aes(x=Enrolment, y= count, fill = Condition)) + 
    geom_bar(position="dodge", stat="identity") +
  labs(title="Enrollment Condition",
        x="Year", y= "Number of Students") +
  geom_text(aes(label = count), 
            position = position_dodge(0.9),
            color="black",vjust = 1.1,hjust = 0.5) + 
  scale_fill_brewer(palette="Reds") + 
  coord_flip() +
  labs(x="")
```


Courses
=====================================  
Row {data-height=100}
-----------------------------------------------------------------------

### Total Number of Courses

```{r}
CourseCount <- courses %>%
  summarise(TotalCourses = n())

  valueBox(value = CourseCount, icon = "fa-pencil", color = "red")
```

### Number of Mandatory Courses

```{r}
MCourseCount <- courses %>%
  filter(Type == "Mandatory") %>%
  summarise(n())

  valueBox(value = MCourseCount, icon = "fa-pencil", color = "red")
```

### Number of Elective Courses

```{r}
ECourseCount <- courses %>%
  filter(Type == "Elective") %>%
  summarise(n())

  valueBox(value = ECourseCount, icon = "fa-pencil", color = "red")
```

Row {data-height=500}
-------------------------------------
### Average Grade by Course 

```{r}
left_join(grades, courses %>% select(Course_ID, Course)) %>%
  group_by(Course) %>% 
  summarize(N = n(), Avg_Grade = mean(Grade)) %>%
  ggplot(aes(x= reorder(Course, -Avg_Grade), y= Avg_Grade, fill = as.factor(Course)), show.legend = FALSE) + 
  labs(title="Average Grade in each Course",
        x="", y= "Grades (Average)") +
  geom_bar(stat = "Identity") + 
  coord_flip() + 
  scale_fill_brewer(palette="Reds") +
  theme(legend.position = "none") 
```


### Number of Students by Course 

```{r}
left_join(grades, courses) %>%
  mutate(passed = ifelse(Grade >= 5,1,0)) %>%
  group_by(Course,Year) %>%
  summarise(students = n(),passed = sum(passed)) %>%
  mutate(frac_passed = passed/students) %>%
  ggplot(aes(y = students,x = Year, fill = Course)) +
  geom_bar(position="dodge", stat="identity", height = 1) +
  coord_flip() +
  scale_fill_brewer(palette="Reds") + 
  labs(x="") 
  

  


```

### Fraction of Students who passed by Course 

```{r}
left_join(grades, courses) %>%
  mutate(passed = ifelse(Grade >= 5,1,0)) %>%
  group_by(Course,Year) %>%
  summarise(students = n(),passed = sum(passed)) %>%
  mutate(frac_passed = passed/students) %>%
  ggplot(aes(y = frac_passed,x = Year, fill = Course)) +
  geom_bar(position="dodge", stat="identity", height = 1) +
  coord_flip() +
  scale_fill_brewer(palette="Reds") +
  labs(x="") 
  
  

  


```

Grades
=====================================  

Row {data-height=100}
-----------------------------------------------------------------------

### Number of Passing Marks

```{r}
CountPassingMarks <- grades %>%
  filter(Grade >= 5) %>%
  summarise(n())

  valueBox(value = CountPassingMarks, color = "red")
```

### Number of Failing Marks

```{r}
CountFailingMarks <- grades %>%
  filter(Grade < 5) %>%
  summarise(n())

  valueBox(value = CountFailingMarks, color = "red")
```

### % Passing Marks

```{r}
PercPassing <- round(CountPassingMarks/(CountPassingMarks + CountFailingMarks) * 100,0)

  valueBox(value = PercPassing, color = "red")
```

Row {data-height=300}
-----------------------------------------------------------------------

### Average Grade in each Course

```{r}
left_join(grades, courses %>% select(Course_ID, Course)) %>%
  group_by(Course) %>% 
  summarize(N = n(), Avg_Grade = mean(Grade)) %>%
  ggplot(aes(x= reorder(Course, -Avg_Grade), y= Avg_Grade, fill = as.factor(Course))) + 
  labs(x="Course", y= "Grades (Average)") +
  geom_bar(stat = "Identity") + 
  coord_flip() + 
  scale_fill_brewer(palette="Reds") +
  theme(legend.position = "none") +
  labs(x = "")
```


### Average Grade per Year 


```{r}
grades %>% group_by(Year) %>% 
  summarize(N = n(), Avg_Grade = mean(Grade)) %>%
ggplot(aes(x=Year, y= Avg_Grade, fill = as.factor(Year))) + 
    geom_bar(position="dodge", stat="identity") +
  labs(x="Year", y= "Average Grade") + 
  scale_y_continuous(n.breaks = 8) +
  scale_fill_brewer(palette="Reds") +
  theme(legend.position = "none") +
  labs(x = "")
```

Row {data-height=300}
-----------------------------------------------------------------------

### Average Grade per Student

```{r}
# Define the number of colors 
nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(9, "Reds"))(nb.cols)

grades %>% group_by(Student_ID) %>% 
  summarize(N = n(), Avg_Grade = mean(Grade)) %>%
ggplot(aes(x=as.factor(Student_ID), y= Avg_Grade, fill = as.factor(Student_ID))) + 
    geom_bar(position="dodge", stat="identity") +
  labs(x="Student ID", y= "Average Grade") + 
    coord_flip() + 
  scale_fill_manual(values = mycolors) +
  theme(legend.position = "none") 


```

### Grade by Gender

```{r}
left_join(grades,students) %>%
  group_by(Gender,Year) %>%
  summarise(AvgGrade = mean(Grade),) %>%
  ggplot(aes(x=Year,y=AvgGrade,fill=Gender)) +
  geom_bar(position = "dodge",stat ="identity") +
  scale_fill_brewer(palette="Reds") +
  labs(x="")

```

### Grade by Nationality

```{r}
nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(9, "Reds"))(nb.cols)
left_join(grades, students) %>%
 group_by(Country) %>% 
  summarize(N = n(), Avg_Grade = mean(Grade)) %>%
ggplot(aes(x= Country, y= Avg_Grade, fill = as.factor(Country))) + 
    geom_bar(position="dodge", stat="identity") +
  labs(x="Country", y= "Average Grade") + 
  scale_y_continuous(n.breaks = 8) +
coord_flip() + 
  scale_fill_manual(values = mycolors) +
  theme(legend.position = "none")

```

Students
=====================================  

Row {data-height=100}
-----------------------------------------------------------------------

### Avg Age of Students

```{r}
  

  AvgAge <- students %>%
    summarise(avg = mean(Age))

  AvgAge <- c(AvgAge)

  valueBox(value = AvgAge, color = "red")
```

### Number of EU Students

```{r}
  NumEU <- students %>%
    filter(Nationality == "EU") %>%
    summarise(n())

  valueBox(value = NumEU, color = "red")
```

### Number of Non-EU Students

```{r}
NumNonEU <- students %>%
    filter(Nationality == "NONEU") %>%
    summarise(n())

  valueBox(value = NumNonEU, color = "red")
```

Row {data-height=300}
-----------------------------------------------------------------------

### Students Gender


```{r}
students %>% group_by(Gender) %>% summarise(N= n()) %>%
  ggplot(aes(x="", y=N, fill=Gender)) +
  geom_bar(stat="identity", width=1, color="white") +
  geom_text(aes(label = N), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  scale_fill_brewer(palette="Reds") +
  theme_void()
```

### Student Admission by Year/Term


```{r}
students %>%
  group_by(Admission.Year,Admission.Month) %>%
  summarise(Count01 = n()) %>%
  ggplot(aes(x = Admission.Year,y=Count01,fill=Admission.Month)) +
  geom_bar(position = "dodge",stat ="identity", width = 1) +
  labs(y="# of Admissions") +
  scale_fill_brewer(palette="Reds") +
  labs(x="")
  
```

Row {data-height=300}
-----------------------------------------------------------------------

### Students Nationality

```{r}
students %>% group_by(Nationality) %>% summarise(N= n()) %>%
  ggplot(aes(x="", y=N, fill=Nationality)) +
  geom_bar(stat="identity", width=1, color="white") +
  geom_text(aes(label = N), position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette="Reds") +
  theme_void()


```

### Students Country

```{r}
# Define the number of colors 
nb.cols <- 20
mycolors <- colorRampPalette(brewer.pal(9, "Reds"))(nb.cols)

students %>% group_by(Country) %>% summarise(N= n()) %>%
  ggplot(aes(x="", y=N, fill=Country)) +
  geom_bar(stat="identity", width=1) +
  geom_text(aes(label = N), position = position_stack(vjust = 0.5)) +
  coord_polar("y", start=0) +
  scale_fill_manual(values = mycolors) +
  theme_void()


```
